# Introduction {#intro}

You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].

Below is an example borrowed from [Petersen](https://www.kellogg.northwestern.edu/faculty/petersen/htm/papers/se/test_data.htm).

```{r essaiCL}
library(sandwich)
## Petersen's data
data("PetersenCL", package = "sandwich")
m <- lm(y ~ x, data = PetersenCL)

## clustered covariances
## one-way
vcovCL(m, cluster = ~ firm)
vcovCL(m, cluster = PetersenCL$firm) ## same
## one-way with HC2
vcovCL(m, cluster = ~ firm, type = "HC2")
## two-way
vcovCL(m, cluster = ~ firm + year)
vcovCL(m, cluster = PetersenCL[, c("firm", "year")]) ## same

```

XXXX

Sargan-Hansen () test. @Sargan_1958 and @Hansen_1982

Durbin-Wu-Hausman test: @Durbin_1954 / @Wu_1973 / @Hausman_1978

[Use R!](https://www.hds.utc.fr/~tdenoeux/dokuwiki/_media/en/kleibzeil_-_aer.pdf) is an excellent tutorial.
(notably for `plm` and the Arellano-Bond example, 140 UK firms)

Program evaluation (very good survey): @Abadie_Cattaneo_2018
Mostly harmless: @angrist_mostly_2008

Diff-in-Diff: [Card and Krueger (1994)](https://towardsdatascience.com/analyze-causal-effect-using-diff-in-diff-model-85e07b17e7b7)

XXXX


